other functionalities

### 2024-02-19 inst/gitbook/05_other-functionalities.Rmd
added manual entry for set.root.time

>>>>

## `set.root.time`

This function can be used to easily add a `$root.time` element to `"phylo"` or `"multiPhylo"` objects.
This `$root.time` element is used by `dispRity` and several packages (e.g. `Claddis` and `paleotree`) to scale the branch length units of a tree allowing them to be usually expressed in million of years (Mya).

For example, on a standard random tree, no `$root.time` exist so the edge lengths are not expressed in any specific unit:

```{r}
## A random tree with no root.time
my_tree <- rtree(10)
my_tree$root.time # is NULL
```

You can add a root time by either manually setting it:

```{r}
## Adding an arbitrary root time
my_tree_arbitrary <- my_tree
## Setting the age of the root to 42
my_tree_arbitrary$root.time <- 42
```

Or by calculating it automatically from the cumulated branch length information (making the youngest tip age 0 and the oldest the total age/depth of the tree)

```{r}
## Calculating the root time from the present
my_tree_aged <- my_tree 
my_tree_aged <- set.root.time(my_tree)
```

If you want the youngest tip to not be of age 0, you can define an arbitrary age for it and recalculate the age of the root from there using the `present` argument (say the youngest tip is 42 Mya old):

```{r}
## Caculating the root time from 42 Mya 
my_tree_age <- set.root.time(my_tree, present = 42)
```

This function also works with a distribution of trees (`"multiPhylo"`).
